Generative Dynamics of Supreme Court Citations: Analysis with a New Statistical Network Model

### 2020-09-02 Tex/PA_appendix_rr/appendix.tex
final revisions

\section{c-ERGM Estimation}

The normalizing constant in Equation 1 of the main text is intractable. For example, in the simple case of adding three cases to a network in which six cases already exists---like that depicted in Figure 4 of the main text---there are 16,777,216 unique configurations of $C_t$ that could be observed. The typical Supreme Court term involves adding hundreds of cases to a network that already includes thousands of previous cases. This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with the c-ERGM.

@@

%We estimate the MCMLE for every term of the c-ERGM between 1950 and 2015. While the network for the 1950 term consists of $1962$ nodes, the size of the network monotonically increases to $10,020$ nodes for the 2015 term, making estimation more challenging as the network grows. The growth in the network also serves as our reason for estimating a separate model for each term. First, the network is large, with each term introducing the potential for hundreds-of-thousands of new citations. This means we have the statistical power necessary to estimate a separate set of parameter values for each term. Second, based on the findings of \citet{shalizi2013consistency}, we know that the set of parameter values that best fits a network should change as more nodes are added to the network. This means it would be inappropriate to assume a constant set of parameter values as more cases are added each term.

The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant. However, the estimation of most networks in the 90s with the MPLE as starting values was not feasible in an reasonable time frame anymore. Instead, we improved the choice of starting value $\theta_0$ by fixing it at the MCMLE of the previous term $t-1$ and successfully obtained the MCMLE of the network at term $t$. But even this approach started to fail for networks around the turn of the millennium. Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}. 
The MCMLE for these large citation networks was obtained by setting the starting value according to an novel approach introduced by \citet{SchmidHunter2020}. This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$. However, the MPLE of networks with the same sufficient statistics is not necessarily the same. Instead of starting the MCMLE algorithm at the network's MPLE, \citet{SchmidHunter2020} propose searching for a new network $C_t^*$ on the same nodes as the observed network that satisfies $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$, and has a weak dependence structure among unfixed ties. Such a network can be found using simulated annealing algorithms \citep{Kirkpatrick83}. For networks with a weak dependence structure among unfixed ties, the MPLE is similar to the MLE, in addition, the same sufficient statistic between $C_t^*$ and the observed network $C_t$ guarantees the same MLE between these two networks. This makes the MPLE of $C_t^*$ an effective starting value for the MCMLE algorithm. Since for some networks the MCMLE was only obtainable using simulated annealing method to find a starting value, the final results in the paper have all been estimated using simulated annealing.
The simulated annealing algorithm for finding an improved starting value for c-ERGMs was implemented in the \textbf{cERGM}-package for \textbf{R} \citep{RCore} and can be found at \url{http://github.com/schmid86/cERGM}.

>>>>

\section{cERGM Estimation}

The normalizing constant in Equation 1 of the main text is intractable. For example, in the simple case of adding three cases to a network in which six cases already exists---like that depicted in Figure 4 of the main text---there are 16,777,216 unique configurations of $C_t$ that could be observed. The typical Supreme Court term involves adding hundreds of cases to a network that already includes thousands of previous cases. This means that straightforward methods of maximum likelihood estimation (MLE) are infeasible with the cERGM.

@@

%We estimate the MCMLE for every term of the cERGM between 1950 and 2015. While the network for the 1950 term consists of $1962$ nodes, the size of the network monotonically increases to $10,020$ nodes for the 2015 term, making estimation more challenging as the network grows. The growth in the network also serves as our reason for estimating a separate model for each term. First, the network is large, with each term introducing the potential for hundreds-of-thousands of new citations. This means we have the statistical power necessary to estimate a separate set of parameter values for each term. Second, based on the findings of \citet{shalizi2013consistency}, we know that the set of parameter values that best fits a network should change as more nodes are added to the network. This means it would be inappropriate to assume a constant set of parameter values as more cases are added each term.

The MCMLE of networks up until the 90s was obtainable in a reasonable time frame starting at the MPLE and sampling $10,000$ networks to approximate the normalizing constant. However, the estimation of most networks in the 90s with the MPLE as starting values was not feasible in an reasonable time frame anymore. Instead, we improved the choice of starting value $\theta_0$ by fixing it at the MCMLE of the previous term $t-1$ and successfully obtained the MCMLE of the network at term $t$. But even this approach started to fail for networks around the turn of the millennium. Neither the MPLE nor the MCMLE of previous terms as starting values led to successful estimation, and neither did the Stepping algorithm \citep{hummel2012improving}. 
The MCMLE for these large citation networks was obtained by setting the starting value according to an novel approach introduced by \citet{SchmidHunter2020}. This method is based on the fact that the MLE of exponential family distributions is solely a function of the vector of sufficient statistics $h(C_t, C_{<t})$, meaning that the MLE of two networks $A$ and $B$ is equal if $h(A)=h(B)$. However, the MPLE of networks with the same sufficient statistics is not necessarily the same. Instead of starting the MCMLE algorithm at the network's MPLE, \citet{SchmidHunter2020} propose searching for a new network $C_t^*$ on the same nodes as the observed network that satisfies $h(C_t, C_{<t}) = h(C_t^*, C_{<t})$, and has a weak dependence structure among unfixed ties. Such a network can be found using simulated annealing algorithms \citep{Kirkpatrick83}. For networks with a weak dependence structure among unfixed ties, the MPLE is similar to the MLE, in addition, the same sufficient statistic between $C_t^*$ and the observed network $C_t$ guarantees the same MLE between these two networks. This makes the MPLE of $C_t^*$ an effective starting value for the MCMLE algorithm. Since for some networks the MCMLE was only obtainable using simulated annealing method to find a starting value, the final results in the paper have all been estimated using simulated annealing.
The simulated annealing algorithm for finding an improved starting value for cERGMs was implemented in the \textbf{cERGM}-package for \textbf{R} \citep{RCore} and can be found at \url{http://github.com/schmid86/cERGM}.

### 2020-09-02 Tex/PA_rr/main.tex
final revisions

The first network property that we theorize in the context of Supreme Court citations is transitivity. In a network of directed relations (e.g., A cites B, but B doesn't cite A) transitivity refers to the tendency for A to send a tie to C if A sends a tie to B and B sends a tie to C \citep{holland1971transitivity}. In undirected networks, transitivity is simply the process by which friends of friends become friends (i.e., a friend of a friend is a friend). The term, ``transitive closure'' refers to a tie forming from A to C in response to extant ties from A to B and B to C. When writing opinions, Supreme Court justices present the legal bases for their rulings, which often involves discussing the most primary/relevant precedents underpinning these legal bases, but also the precedents and legal rules on which the primary precedents were based. This process of presenting several layers/levels of precedent in an opinion follows the structure of transitive closure exactly---opinion A cites opinion B as a primary precedent, and then cites opinion C because opinion B cites opinion C. The two examples presented below illustrate this process.

@@

We develop the citation ERGM (c-ERGM), to account for the structural constraints that apply to the network of Supreme Court citations. These structural constraints amount to three departures from the structure of networks for which ERGMs are currently designed. First, the citation network is partially acyclic. If two cases are decided during the same term, they can cite each other, forming a mutual edge (or two-cycle). However, if case $i$ is decided before case $j$, case $j$ can cite case $i$, but case $i$ cannot cite case $j$. Second, new edges can be created over time, but cannot be eliminated. Unlike in, e.g., an alliance network, in which two countries can dissolve an alliance, once a citation exists in a citation network it cannot be dissolved. Third, the set of nodes in the network must increase for new edges to be created. In conventional ERGMs, the number of nodes in the network can increase or decrease in each time period, and is typically stable over time. In a citation network, new edges (and non-edges) are introduced over time via the introduction of new nodes. The structure of the citation network is depicted in Figure \ref{fig:ctergm}---a hypothetical citation network established over three time periods, with three cases decided in each time period. We denote $C_t$ to be the set of citation and non-citations added to the network at time $t$ (i.e., via the addition of three cases), $C_{ <t}$ to be the citations among cases decided before time $t$, and $C_{ \leq t}$ to denote the entire set of citations and non-citations on which $C_t$ can depend through the c-ERGM specification.

@@

The likelihood function of the c-ERGM is given by
\begin{equation}
l(\bm{\theta},C_{\leq t}) =  \frac{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}) } \right] }{ \sum_{C_t^* \in \mathcal{C}_t} \exp \left[ {\bm{\theta}'\bm{h}(C^*_{t},C_{<t}) }\right]  },
\end{equation}
where $\bm{\theta}$ is a vector of real-valued parameters, and $\bm{h}(C_{t},C_{<t})$ is a vector of scalar-valued functions that each quantify a feature of the citation network (e.g., the relationship between citation ties and a case attribute, the number of mutual edges in the citation network). $\exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}) } \right]$ is a positive weight that is proportional to the probability of observing any particular form of the citations and non-citations added to the network at time $t$.  The denominator of Equation 1 represents a normalizing constant, in which the positive weight is summed over all possible configurations of $C_{t}$, from the network in which the cases added to the network at time $t$ send no citations at all, to the network in which the cases added at time $t$ cite every possible case, and everything in between. %The probability of observing the citation network up to time $T$---the final time point in the dataset---is given by the product over the probability of adding the citations at each time point $t$, conditional on the citations that were added to the network previously.

Though the likelihood function of the c-ERGM may appear quite different from that of statistical models conventionally used in political science, analysis of the conditional probability of a single citation from case $i$ to case $j$ reveals that we can interpret the parameters similar to logistic regression coefficients. $$ P(C_{ij,t} = 1 | C_{-ij,t}, C_{ < t}) = \frac{\exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right]}{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right] + \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 0) } \right]}, $$ $$ \text{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~} = \frac{1}{ 1 + \exp \left[ - {\bm{\theta}'\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)} \right]}, $$
where $C_{ij,t} = 1$ indicates that case $i$ cites case $j$, $C_{ij,t} = 0$ indicates that case $i$ does not cite case $j$, $C_{-ij,t}$ is the observed elements of $C_{t}$ except $C_{ij,t}$, and $\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)$ is the change in $\bm{h}(C_{t},C_{<t})$ that results from toggling $C_{ij,t} = 0$ to $C_{ij,t} = 1$. This re-arrangement illustrates that the parameters can be interpreted in terms of the change in the log odds of a citation from $i$ to $j$ given a one-unit increase in the corresponding element of $\bm{h}$, conditional on the other citations observed in the network. For example, if the value of $\theta$ corresponding to an element of $\bm{h}$ that counts the number of mutual edges in the network is 0.5, then the log odds of observing $C_{ij,t} = 1$ increases by 0.5 if case $i$ is cited by case $j$ (as compared to the configuration in which case $j$ does not cite case $i$). The logit form conditional probability is well known for the ERGM family \citep{goodreau2009birds}.

@@

We fit a c-ERGM for every Supreme Court term between 1950 and 2015, meaning that we have 66 models where the outcome variables are the citations sent during the given term. Our approach of fitting separate term-by-term models is motivated by the temporally-dynamic data generating process. First, as our results will show, the temporal context and justice composition of a court will lead to differences in how opinions are written and therefore how citations are used. Second, based on the findings of \citet{shalizi2013consistency}, we know that the set of parameter values that best fits a network should change as more nodes are added to the network. This means it would be inappropriate to assume a constant set of parameter values as more cases are added each term. Practically, our dataset is large enough to support this approach, with each term introducing the potential for hundreds-of-thousands of new citations. This gives us the statistical power necessary to estimate a separate set of parameter values for each term. Further, while the entire set of citations can in theory be fit in a single network with varying parameters, it is technically difficult. While the network for the 1950 term consists of a manageable $1,962$ nodes, the size of the network monotonically increases to $10,020$ nodes in the 2015 term, making estimation more challenging as the network grows.

Estimation of the c-ERGM parameters, which is done with a computationally-intensive, simulation-based approach, is presented in detail in the appendix. The basis of our computation is the \textbf{ergm}-package \citep{HunterDavidR..2008} in \textbf{R} \citep{RCore}. In addition, we developed a wrapper \textbf{R}-package called \textbf{cERGM} \citep{cergm} to conveniently fit citation network data using the c-ERGM.

@@

\subsection{c-ERGM specification}

The c-ERGM we estimate includes two classes of terms---one that captures the effects of covariates on tie formation, and another that captures the complex dependence processes that we expect to observe in the Supreme Court Citation Network. We describe our model specification by defining the terms within these two classes.

@@

Covariate effects are accounted for in the c-ERGM via the term that is used to specify the effect of covariates in other ERGM family models, as
$$h_{covariate}(C_t,C_{<t},X) =  \sum_{ij} C_{ij,t}X_{ij}.$$ Since $C_{ij,t}$ is a binary indicator of whether case $i$ cites case $j$, $h_{covariate}$ amounts to the sum of covariate values among directed dyads for which we observed a citation. The dyadic interpretation of the coefficient attributed to this term is the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in $X_{ij}$ (i.e., exactly the interpretation of the effect of a covariate in logistic regression). We include several exogenous covariates based on this standard formulation.

@@

\subsection{c-ERGM Coefficient Estimates}
The estimates from the c-ERGM are presented in Figures \ref{SCC_results_1} and \ref{SCC_results_2}. In these figures we depict coefficient estimates and 95\% confidence intervals for all of the effects for all the variables we discussed in sections \ref{covariate_terms} and \ref{dependence_terms}. The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in the respective variable. In our figures, a circle indicates that the estimate was significant at an $\alpha=0.05$ level, while an square translates to a p-value between $0.05$ and $0.1$. A triangle signals that the estimate for a given year was not significant at an $\alpha=0.1$ level. In each panel, the background shading indicates different chief justice courts, with the two darker shades representing, respectively, the Warren (1954-1969) and Rehnquist (1987-2005) courts. Finally, note that the panels showing results of the reciprocity statistic and the \textit{Overruled Cases} variable do not provide estimates for every term. In these missing terms there were no two cases that cited each other or no overruled cases being cited, making the estimate of these terms equivalent to $-\infty$. %For those years, the reciprocity term is included in the model, but the estimate of $-\infty$.

@@

\subsubsection{Covariate Terms}
Effects of the exogenous covariates included in the c-ERGM are presented in Figure \ref{SCC_results_2}. Most of the exogenous covariate effects align with expectations and have relatively stable parameter estimates over time. We find that cases are more likely to cite those that (1) have been decided most recently, (2) were authored by the same justice, (3) are in the same issue area, and (4) have not been overruled. Surprisingly, we do not find consistent effects for majority coalition size or for the two covariates that involve justice ideology. The effect of majority coalition size was generally not statistically significant until the later portion of the period examined; ideological distance between two cases had signs and significance levels that are highly inconsistent; and the effect of the ideological breadth of the potentially cited case was effectively zero for the entire period.

@@

In the above discussion, we identified an explanation for the observed temporal dynamics in the effect of ideological distance. Specifically, ideologically distant cases become more important as courts adopt the relatively new `split-the-difference' jurisprudential approach which requires broad citations. We recognize that there are alternative explanations---perhaps one that is related to the coterminous change in the effect of majority coalition size---but a deeper look into the relevant literature is beyond the scope of this paper. Our discussion serves to illustrate that term-by-term c-ERGMs can uncover important temporal dynamics that courts researchers can use to study a much greater range of citation-related phenomenon beyond what we present here. To the extent that the US Supreme Court is a political entity that interacts with the broader society, the body of court opinions and citations between them is corpus that continues to grow based on both its history and its contemporary context. The methods we present here afford researchers the tools to better understand these complexities.

@@

We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network. This methodology---the citation-ERGM---enables researchers to include both exogenous covariates such as the ideological predisposition and age of a case, and dependence terms, such as transitivity and reciprocity, as explanations for citation formation. We apply this methodology to a network that includes all Supreme Court cases decided between 1950 and 2015. We find, somewhat counterintuitively, that Supreme Court citations are highly reciprocal. We also find that citations are driven by dependencies such as triad closure and popularity. The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model. The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.

>>>>

The first network property that we theorize in the context of Supreme Court citations is transitivity. In a network of directed relations (e.g., $i$ cites $j$, but $j$ doesn't cite $i$) transitivity refers to the tendency for $i$ to send a tie to $k$ if $i$ sends a tie to $j$ and $j$ sends a tie to $k$ \citep{holland1971transitivity}. In undirected networks, transitivity is simply the process by which friends of friends become friends (i.e., a friend of a friend is a friend). The term, ``transitive closure'' refers to a tie forming from $i$ to $k$ in response to extant ties from $i$ to $j$ and $j$ to $k$. When writing opinions, Supreme Court justices present the legal bases for their rulings, which often involves discussing the most primary/relevant precedents underpinning these legal bases, but also the precedents and legal rules on which the primary precedents were based. This process of presenting several layers/levels of precedent in an opinion follows the structure of transitive closure exactly---opinion $i$ cites opinion $j$ as a primary precedent, and then cites opinion $k$ because opinion $j$ cites opinion $k$. The two examples presented below illustrate this process.

@@

We develop the citation ERGM (cERGM), to account for the structural constraints that apply to the network of Supreme Court citations. These structural constraints amount to three departures from the structure of networks for which ERGMs are currently designed. First, the citation network is partially acyclic. If two cases are decided during the same term, they can cite each other, forming a mutual edge (or two-cycle). However, if case $i$ is decided before case $j$, case $j$ can cite case $i$, but case $i$ cannot cite case $j$. Second, new edges can be created over time, but cannot be eliminated. Unlike in, e.g., an alliance network, in which two countries can dissolve an alliance, once a citation exists in a citation network it cannot be dissolved. Third, the set of nodes in the network must increase for new edges to be created. In conventional ERGMs, the number of nodes in the network can increase or decrease in each time period, and is typically stable over time. In a citation network, new edges (and non-edges) are introduced over time via the introduction of new nodes. The structure of the citation network is depicted in Figure \ref{fig:ctergm}---a hypothetical citation network established over three time periods, with three cases decided in each time period. We denote $C_t$ to be the set of citation and non-citations added to the network at time $t$ (i.e., via the addition of three cases), $C_{ <t}$ to be the citations among cases decided before time $t$, and $C_{ \leq t}$ to denote the entire set of citations and non-citations on which $C_t$ can depend through the cERGM specification.

@@

The likelihood function of the cERGM is given by
\begin{equation}
l(\bm{\theta},C_{\leq t}) =  \frac{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}) } \right] }{ \sum_{C_t^* \in \mathcal{C}_t} \exp \left[ {\bm{\theta}'\bm{h}(C^*_{t},C_{<t}) }\right]  },
\end{equation}
where $\bm{\theta}$ is a vector of real-valued parameters, and $\bm{h}(C_{t},C_{<t})$ is a vector of scalar-valued functions that each quantify a feature of the citation network (e.g., the relationship between citation ties and a case attribute, the number of mutual edges in the citation network). $\exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}) } \right]$ is a positive weight that is proportional to the probability of observing any particular form of the citations and non-citations added to the network at time $t$.  The denominator of Equation 1 represents a normalizing constant, in which the positive weight is summed over all possible configurations of $C_{t}$, from the network in which the cases added to the network at time $t$ send no citations at all, to the network in which the cases added at time $t$ cite every possible case, and everything in between. %The probability of observing the citation network up to time $T$---the final time point in the dataset---is given by the product over the probability of adding the citations at each time point $t$, conditional on the citations that were added to the network previously.

Though the likelihood function of the cERGM may appear quite different from that of statistical models conventionally used in political science, analysis of the conditional probability of a single citation from case $i$ to case $j$ reveals that we can interpret the parameters similar to logistic regression coefficients. $$ P(C_{ij,t} = 1 | C_{-ij,t}, C_{ < t}) = \frac{\exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right]}{ \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) } \right] + \exp \left[ {\bm{\theta}'\bm{h}(C_{t},C_{<t}| C_{ij,t} = 0) } \right]}, $$ $$ \text{~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~} = \frac{1}{ 1 + \exp \left[ - {\bm{\theta}'\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)} \right]}, $$
where $C_{ij,t} = 1$ indicates that case $i$ cites case $j$, $C_{ij,t} = 0$ indicates that case $i$ does not cite case $j$, $C_{-ij,t}$ is the observed elements of $C_{t}$ except $C_{ij,t}$, and $\left(\bm{h}(C_{t},C_{<t}| C_{ij,t} = 1) - \bm{h}(C_{t},C_{<t}| C_{ij,t} = 0)\right)$ is the change in $\bm{h}(C_{t},C_{<t})$ that results from toggling $C_{ij,t} = 0$ to $C_{ij,t} = 1$. This re-arrangement illustrates that the parameters can be interpreted in terms of the change in the log odds of a citation from $i$ to $j$ given a one-unit increase in the corresponding element of $\bm{h}$, conditional on the other citations observed in the network. For example, if the value of $\theta$ corresponding to an element of $\bm{h}$ that counts the number of mutual edges in the network is 0.5, then the log odds of observing $C_{ij,t} = 1$ increases by 0.5 if case $i$ is cited by case $j$ (as compared to the configuration in which case $j$ does not cite case $i$). The logit form conditional probability is well known for the ERGM family \citep{goodreau2009birds}.

@@

We fit a cERGM for every Supreme Court term between 1950 and 2015, meaning that we have 66 models where the outcome variables are the citations sent during the given term. Our approach of fitting separate term-by-term models is motivated by the temporally-dynamic data generating process. First, as our results will show, the temporal context and justice composition of a court will lead to differences in how opinions are written and therefore how citations are used. Second, based on the findings of \citet{shalizi2013consistency}, we know that the set of parameter values that best fits a network should change as more nodes are added to the network. This means it would be inappropriate to assume a constant set of parameter values as more cases are added each term. Practically, our dataset is large enough to support this approach, with each term introducing the potential for hundreds-of-thousands of new citations. This gives us the statistical power necessary to estimate a separate set of parameter values for each term. Further, while the entire set of citations can in theory be fit in a single network with varying parameters, it is technically difficult. While the network for the 1950 term consists of a manageable $1,962$ nodes, the size of the network monotonically increases to $10,020$ nodes in the 2015 term, making estimation more challenging as the network grows.

Estimation of the cERGM parameters, which is done with a computationally-intensive, simulation-based approach, is presented in detail in the appendix. The basis of our computation is the \textbf{ergm}-package \citep{HunterDavidR..2008} in \textbf{R} \citep{RCore}. In addition, we developed a wrapper \textbf{R}-package called \textbf{cERGM} \citep{cergm} to conveniently fit citation network data using the cERGM.

@@

\subsection{cERGM specification}

The cERGM we estimate includes two classes of terms---one that captures the effects of covariates on tie formation, and another that captures the complex dependence processes that we expect to observe in the Supreme Court Citation Network. We describe our model specification by defining the terms within these two classes.

@@

Covariate effects are accounted for in the cERGM via the term that is used to specify the effect of covariates in other ERGM family models, as
$$h_{covariate}(C_t,C_{<t},X) =  \sum_{ij} C_{ij,t}X_{ij}.$$ Since $C_{ij,t}$ is a binary indicator of whether case $i$ cites case $j$, $h_{covariate}$ amounts to the sum of covariate values among directed dyads for which we observed a citation. The dyadic interpretation of the coefficient attributed to this term is the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in $X_{ij}$ (i.e., exactly the interpretation of the effect of a covariate in logistic regression). We include several exogenous covariates based on this standard formulation.

@@

\subsection{cERGM Coefficient Estimates}
The estimates from the cERGM are presented in Figures \ref{SCC_results_1} and \ref{SCC_results_2}. In these figures we depict coefficient estimates and 95\% confidence intervals for all of the effects for all the variables we discussed in section \ref{covariate_terms}. The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one-unit increase in the respective variable. In our figures, a circle indicates that the estimate was significant at an $\alpha=0.05$ level, while an square translates to a p-value between $0.05$ and $0.1$. A triangle signals that the estimate for a given year was not significant at an $\alpha=0.1$ level. In each panel, the background shading indicates different chief justice courts, with the two darker shades representing, respectively, the Warren (1954-1969) and Rehnquist (1987-2005) courts. Finally, note that the panels showing results of the reciprocity statistic and the \textit{Overruled Cases} variable do not provide estimates for every term. In these missing terms there were no two cases that cited each other or no overruled cases being cited, making the estimate of these terms equivalent to $-\infty$. %For those years, the reciprocity term is included in the model, but the estimate of $-\infty$.

@@

\subsubsection{Covariate Terms}
Effects of the exogenous covariates included in the cERGM are presented in Figure \ref{SCC_results_2}. Most of the exogenous covariate effects align with expectations and have relatively stable parameter estimates over time. We find that cases are more likely to cite those that (1) have been decided most recently, (2) were authored by the same justice, (3) are in the same issue area, and (4) have not been overruled. Surprisingly, we do not find consistent effects for majority coalition size or for the two covariates that involve justice ideology. The effect of majority coalition size was generally not statistically significant until the later portion of the period examined; ideological distance between two cases had signs and significance levels that are highly inconsistent; and the effect of the ideological breadth of the potentially cited case was effectively zero for the entire period.

@@

In the above discussion, we identified an explanation for the observed temporal dynamics in the effect of ideological distance. Specifically, ideologically distant cases become more important as courts adopt the relatively new `split-the-difference' jurisprudential approach which requires broad citations. We recognize that there are alternative explanations---perhaps one that is related to the coterminous change in the effect of majority coalition size---but a deeper look into the relevant literature is beyond the scope of this paper. Our discussion serves to illustrate that term-by-term cERGMs can uncover important temporal dynamics that courts researchers can use to study a much greater range of citation-related phenomenon beyond what we present here. To the extent that the US Supreme Court is a political entity that interacts with the broader society, the body of court opinions and citations between them is corpus that continues to grow based on both its history and its contemporary context. The methods we present here afford researchers the tools to better understand these complexities.

@@

We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network. This methodology---the citation ERGM---enables researchers to include both exogenous covariates such as the ideological predisposition and age of a case, and dependence terms, such as transitivity and reciprocity, as explanations for citation formation. We apply this methodology to a network that includes all Supreme Court cases decided between 1950 and 2015. We find, somewhat counterintuitively, that Supreme Court citations are highly reciprocal. We also find that citations are driven by dependencies such as triad closure and popularity. The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model. The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.
